SoK: Defending Against Web Application Attacks: From Research Approaches to Practical Tools

### 2014-11-14 docs/questioning.tex
1) Changed some section titles. 2) Footnotes became refs. 3) Providing examples in the second section.

For the past several years,
these attacks have been topping the lists of the most dangerous vulnerabilities
published by {\sc owasp}~\cite{OWASPtop10},
{\sc mitre}~\cite{MITREtop25}, and other organizations.
For instance, consider the case of {\sc owasp}'s popular Top Ten
project~\cite{OWASPtop10},
which aims to raise awareness about web application security by
identifying some of the most critical risks organizations may face.
In its three consecutive Top Ten lists (2007, 2010, 2013), different source
code-driven injection attacks dominate the top five positions.
At the same time,
attackers find new ways~\cite{HNSHS12,DKH14} to bypass defense
mechanisms using a variety of techniques, despite the numerous
countermeasures that are being introduced. As an example, already by 2006,
there were more than 20 proposed defenses
against {\sc sql} injection attacks~\cite{HVO06}.
Since then, the number has doubled, while researchers have indicated that
the number of {\sc sql} injection attacks has been steadily
increasing in recent years~\cite{SSL12}.

In this paper, we explore how different attacks associated with 
the exploitation of untrusted input validation errors can be modeled 
under a {\it common perspective}. To that end, we propose an exploitation
model which highlights that most of the steps needed to mount
different types of source code injection attacks are
common.\footnote{Note that we do not consider lower-level attacks based on
the exploitation of memory corruption vulnerabilities and the injection
of binary code---see Section~\ref{sec:attacks}.} This is validated by the fact that
there are protection mechanisms that can defend against more than one of these
types of attacks. Then, we categorize a selection of these protection
mechanisms. In our selection we include approaches that
have been proposed in publications cited more than
20 times. We also include some recent mechanisms that have been
presented in top security conferences, even though they have been
cited less than twenty times, so that recent research is not penalized.
Finally, we analyze each
mechanism across the following dimensions:
\begin{itemize}
\item {\it Accuracy:} protection mechanisms are as good
  as their detection capability; this requires low false positive and
  false negative rates.
\item {\it Availability:} whether the mechanism and its
  testbed are publicly available.
\item {\it Computational performance:} the computational
  overhead of the mechanisms at their points of deployment.
\item {\it Ease of use:} whether the protection
  mechanism is practical in terms of deployment
  and can be easily adopted by security experts.
\item {\it Security:} the robustness of the protection mechanism against
  attackers with knowledge of its internals who attempt to circumvent it.
\item {\it Detection Point:} the location where a mechanism detects an attack
  based on our exploitation model.
\end{itemize}

@@

The rest of the paper is organized as follows.
Section~\ref{sec:attacks} presents our proposed model
of web application attacks based on source code injection.
Section~\ref{sec:dimensions} introduces the dimensions
across which we analyze the various defenses.
Our categorization and analysis is presented in
Section~\ref{sec:defs} and a discussion is provided in
Section~\ref{sec:discussion}. Finally,
Section~\ref{sec:lessons-learned} highlights
some lessons learned from our observations
and Section~\ref{sec:conclusion} concludes the paper.

\section{Attacks}
\label{sec:attacks}

@@

\subsection{Source Code Injection Vulnerabilities}

Figure~\ref{fig:taxonomy} presents a taxonomy of source
code injection attacks. Such attacks may involve source code, either of a
{\it Domain Specific Language} ({\sc dsl}) or a {\it Dynamic Language}.

Application attacks that involve {\sc dsl}s constitute an important
subset of the code injection problem, as {\sc dsl}s like {\sc sql} and
{\sc xml} play an significant role in the development of both web and
mobile applications. For example, many applications have interfaces
where a user enters input to interact with the application's data,
thereby interacting with the underlying database. This input can
become part of an {\sc sql} query and get executed on the target
database. A code injection attack that exploits the vulnerabilities of
these interfaces by taking advantage of input validation issues like
incorrectly passed parameters or incorrect type handling is called an
{\sc sql} injection attack~\cite{CERT02,MS09,HVO06,SW06}. By using
similar techniques malicious users can perform other exploits based on
{\sc dsl}s like {\sc xp}ath~\cite{SW06,CDL07,MKS09}, {\sc
  xml}~\cite{MSM13} and {\sc json}~\cite{SMS13}. With this kind of
attacks, a malicious user can view sensitive information, destroy or
modify protected data, or even crash the entire application. {\sc
  html} is another {\sc dsl} that can be used for malicious purposes
when an application does not properly handle user supplied data. Based
on this vulnerability an attacker can supply valid {\sc html},
typically via a parameter value, and inject their own content into the
page. {\sc html} injection is mainly associated with {\sc xss}
defects~\cite{HNSHS12}.

A recent class of code injection attacks involve dynamic languages
like JavaScript and {\sc php}~\cite{SFVM09,EWKK09,SMS13}. JavaScript
injection attacks, in particular, make up a large subset of dynamic
language-driven attacks. Such attacks are manifested when a web
application accepts and redisplays data of uncertain origin without
appropriate validation and filtering. Based on this flaw, an attacker
can manage to inject a script in the JavaScript engine of a browser
and alter its execution flow~\cite{ELX07}. JavaScript injection
attacks are considered a critical issue in web application security
mainly because they are associated with major vulnerabilities such as
{\sc xss} attacks~\cite{SG07}, {\sc csrf} attacks~\cite{BJM08,LZRL09}
and Cross-Channel Scripting ({\sc xcs}) attacks~\cite{W10,BBB09}. In
many cases {\sc xss} attacks involve the injection of both {\sc html}
and JavaScript code.

% {\sc html} injection --- such attacks pose no threat to cookies and session IDs.

\subsection{Exploitation Model}
\label{sec:model}

In this section we provide a step-by-step exploitation model to
understand the process of carrying out application attacks based on
source code injection vulnerabilities. Figure~\ref{fig:attacks}
presents the steps of different application attacks. Links between
steps are labelled with the paths constituting different attacks. The
sequence of steps for each attack is shown by the numbers next to the
attack names in the labels. Most steps are common in all attacks. This
is because they are based on similar attack vectors as we mentioned in
the previous section. Large crosses on transitions in
Figure~\ref{fig:attacks} indicate the points where defenses detect
attacks.

@@

Once the injected code reaches the vulnerable application, it becomes
a part of a value represented by a program variable. The target of the
attack determines the route from now on. In {\sc sql} injection and
{\sc csrf} attacks the injected code becomes part of a query that
finally reaches the database where it is executed.

@@

\begin{figure*}
\begin{center}
\leavevmode
\includegraphics[scale=0.61]{attacks-steps-new.pdf}
\end{center}
\caption{\label{fig:attacks}Attack model demonstrating web
application attacks based on source
code injection vulnerabilities. For every arrow there are
corresponding transition numbers for
every attack. Every {\it point} indicates a transition where mechanisms
detect attacks. For every point we provide the corresponding
references:
Point {\sc ub} (at the user's browser):~\cite{KJKV09,LV09,TNH07,NSS06,APKLM10,ML10,YCIS07,PSC09,VDDPJ11,OWVS08,DDHPJ10,VFJKKV07,SLMS14,BV08},
Point {\sc s}t{\sc b} (en route from the server to the browser):~\cite{RDWDE07,JKK06a,GC09,JB07,NLC07,WPLKK09,JEP08,PS11},
Point {\sc e}i{\sc q} (embedded in a query en route to the database):~\cite{BWS05,SW06,HCF05,XBS06,PB05,PMP11,MS09,HO05,SMS13},
Point i{\sc db} (in the database):~\cite{BK04,LLW02,VMV05}.}
\end{figure*}

@@

To ensure replicability of research results the source code
implementing a detection mechanism should be available to researchers.
Ideally the code should be available under an open source license, so
that it is easy to modify and improve an approach; even when this is
not possible, for whatever reason, it is important to make sure that
all computer code and test data is available noting any restrictions
on accessibility. Availability of computer code has been recognised as
an important issue outside the computer science field: the journal
\emph{Nature}, in a recent editorial, adopted a publication policy stressing
access to code and data~\cite{nature2014}. At a time and date that the
merits of open access to code are discussed in non computer science
journals~\cite{easterbrook2014}, we should expect computer scientists
to lead the way.

@@

\begin{figure*} [ht]
\begin{center}
\leavevmode
\includegraphics[scale=0.65]{defenses.pdf}
\end{center}
\caption{\label{fig:defenses}The basic categories of countermeasures
against web application attacks, based on source code
injection. For each approach we provide the references
that present corresponding mechanisms:
{\it Policy Enforcement}:~\cite{NSS06,JKK06a,KKVJ06,KJKV09,TNH07,RDWDE07,YCIS07,OWVS08,PSC09,ML10,DDHPJ10,PS11,VDDPJ11,BV08},
{\sc isr}:~\cite{BK04,JB07,GC09,APKLM10},
{\it Parse-Tree Validation}:~\cite{BWS05,SW06},
{\it Taint Tracking}:~\cite{HCF05,PB05,XBS06,NLC07,VFJKKV07,PMP11,SLMS14},
{\it Training}:~\cite{LLW02,HO05,HO06,HO05b,VMV05,JEP08,WPLKK09,MS09,MKS09,MKLS11},
{\it Hybrid}:~\cite{BV08,LV09,SMS13}.}
\end{figure*}

We categorize and describe the various mechanisms developed to detect
the attacks described in Section~\ref{sec:attacks} on the dimensions
we proposed apart from availability, which we treat separately in the
discussion in Section (\ref{sec:discussion}).

Figure~\ref{fig:defenses} presents a taxonomy of the
defenses that protect web applications against attacks based on
code injection vulnerabilities.
We can identify three approaches:
{\it etiological}, {\it symptomatic}, and {\it hybrid}.
The etiological category involves mechanisms designed to
stop attacks based on their causes and origins~\cite{JL75,L81}. 
The symptomatic category incorporates a variety of schemes that
inspect the behavior of applications and detect attacks based on
their undesirable symptoms~\cite{D76,A00}.
Hybrid mechanisms borrow characteristics from both
categories. Table~\ref{tab:comp} groups the various subcategories and
for every mechanism provides the following:

@@

A number of the frameworks define policies based on information and
features provided by the Document Object Model ({\sc dom}) of a web
page. Specifically, developers must place all legitimate scripts
inside {\sc html} elements like {\tt div}. The web browser (Point {\sc
  ub}), parses the {\sc dom} tree and executes scripts only when they
are contained in such elements. All other scripts are treated
according to the policies defined on the server. Frameworks that
support this functionality are {\sc beep}~\cite{TNH07} and {\sc
  dsi}~\cite{NSS06}. The main problem with these mechanisms is that
they do not examine the script's location inside the web document.
Based on this fact, attackers can perform mimicry attacks~\cite{WS02}.
Specifically, they can execute legitimate scripts, but not as intended
by the original design of the developers. This is extensively
described by {\it Athanasopoulos et al.}~\cite{APKLM10}, who also describe
another recent variation of JavaScript injection attacks, known as
return-to-JavaScript attacks that can be used to bypass the above
mechanisms.
% Such attacks share
% many similarities with the {\it return-to-lib}
% attack~\cite{SPPGMMB04}. 
%% PL Explain what it does, I guess it's what you go on to say
%% directly below?
% When performing an attack like this the malicious user transfers the
% execution to points that contain script code which in turn is used to
% perform the attack.

@@

An issue regarding the above frameworks
involves features like script inclusion
and {\tt iframe} tags. Even though they allow developers
decide if they will disable them or not,
this is unpractical because such features are quite popular.
If developers choose to use them, these frameworks cannot
define policies that restrict the behavior of third-party
scripts introduced by such features. For instance,
consider a web page (foo.net) that prints the value
of a query parameter ({\tt query}) from the
page's {\sc url} in the web page's content
without escaping the value. A malicious user
can take advantage of this and inject an {\tt iframe}
into the page to steal the user's cookie and
send it via an image request to a web site
that he or she controls (malicious.com).
This could be achieved by including the following
link to the malicious web site (or send it via phishing
email) and induce the user to click on it:

\noindent {\it WebJail}~\cite{VDDPJ11}, and {\sc soma}~\cite{OWVS08}
are two frameworks that can actually detect such an attack (Point {\sc
  ub}). To achieve this, {\sc soma} requires site administrators to
specify legitimate, external domains for sending or receiving
information in order to approve interactions between them and the
protected web site. As a result, {\sc soma} can also detect {\sc csrf}
attacks. WebJail contains the functionality of third-party scripts by
introducing a web component integrator that restricts the access that
these scripts may have to either the data or the functionality of
other components.

@@

Defenses based on training include some mechanisms that can be easily
circumvented. For example, {\sc didafit} and {\it libAnomaly} do not
tag transactions with their corresponding call sites. This can lead
easily to false negatives. For instance, consider an application that
will send the password for a forgetful user via email by executing the
following query: \bgroup \lstset{language=SQL}
\begin{small}
\begin{lstlisting}
SELECT password from userdata WHERE id = 'Alice'
\end{lstlisting}
\end{small}
\egroup
\noindent
This same application could allow users to lock their terminal,
but allow the unlocking either with the user's password or with
the administrator password (the 4.3 {\sc bsd} {\em lock} command
behaved in this peculiar way).
The corresponding query to verify the password on the locked
workstation would be as follows:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
SELECT password from userdata WHERE id = 'Alice' OR
id = 'admin'
\end{lstlisting}
\end{small}
\egroup
\noindent
In this case {\sc didafit} and {\it libAnomaly}, could be easily bypassed
since a malicious user could obtain the administrator's password
by email by entering on the password retrieval form
as his user identifier the following:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
nosuchuser' OR id = 'admin
\end{lstlisting}
\end{small}
\egroup
\noindent
The infected query matches the signature of the second one above and
is therefore accepted. The problem lies with the call location of the
query, not the query alone.

@@

\section{Discussion}
\label{sec:discussion}

@@

Regarding the subjects of tests, we see that 27 out of 34 mechanisms
were tested on real-world applications known to be vulnerable. The
vulnerabilities associated with these applications are enlisted in the
following providers: the Common Vulnerabilities and Exposures ({\sc
  cve}) database,\footnote{\url{https://cve.mitre.org/}} the {\it
  Bugtraq}\footnote{\url{http://seclists.org/bugtraq/}} security
mailing list, the {\sc xss}ed.com\footnote{\url{www.xssed.com}}
security bulletin provider, Microsoft's security
bulletin\footnote{\url{https://technet.microsoft.com/en-us/security/bulletin}}
and the ha.ckers.org\footnote{\url{www.xssed.com}} security bulletin
provider. Figure~\ref{fig:defect_sources} presents how many, in percentage,
and which, in the figure's caption, authors referred
to which source. In numerous occassions authors performed tests based
on the same applications. For example, 12 mechanisms were tested on a
vulnerable version of {\sc phpbb}.\footnote{https://www.phpbb.com/}

@@

\section{Lessons Learned}
\label{sec:lessons-learned}

@@

@@

When we introduced specicifity and sensitivity in
Section~\ref{ssec:diagnostic-performance} we deferred discussion of
\textsc{ppv} and \textsc{npv} to this point. These two relate to the
performance of a detection mechanism in an actual production setting,
instead of a testbed. If an attack is detected in a production
environment, how much should we be worried? The answer is provided by
\textsc{ppv}. If no attack is detected in a production environment,
how relaxed should we be that no attack has indeed taken place? The
answer is provided by \textsc{npv}. We can calculate \textsc{ppv} and
\textsc{npv} with the following equations:

@@

Apart from testing practices, another area that merits improvement is
the availability of source code and testbeds. We are not aware of
specific reasons why authors of attack detection mechanisms seem to be
wary of releasing online their code and tests. Our finding may reflect
the status in the current point in time, when authors are urged to
publish their code and tests, and may start, or may have already
started doing so, but this does not show yet in the research we
examined, which goes several years back. We also saw instances where
material was published, but does not seem to be available any more.
That points to the importance of curation of research materials. That
is a problem arising in all scientific fields. It is not enough to
publish the underlying code and data, but to make sure that it remains
accessible and to provide the means to test it even as technology
progresses and operating systems, file formats, and software libraries
change. That may be too much to ask right now; what seems reasonable,
though, is to ask of researchers in attack detection mechanisms not to
buck the trend and to take steps to increase the availability of their
research.

@@

% discuss Availability. Finally,

@@

Despite many approaches that have been developed,
application attacks based on source code injection
have been consistently present for the last 15 years,
and it appears that they will continue to be.
Attackers seem to find new ways to introduce
malicious source code to applications by using a
variety of languages and techniques~\cite{HNSHS12,DKH14}.
Meanwhile, during the last decade, there are
numerous mechanisms that have been developed to
detect one or more of such attacks. However,
even if some of them share characteristics
(for instance, {\sc html} sanitization and {\tt eval} handling)
with supported frameworks like Mozilla's Content Security
Policy\footnote{\url{https://developer.mozilla.org/en-US/docs/Web/Security/CSP}}
({\sc csp}) and Google
Caja\footnote{\url{https://code.google.com/p/google-caja/}}~\cite{M06}
the majority of them are not used in practice.

>>>>

For the past several years,
these attacks have been topping the lists of the most dangerous vulnerabilities
published by {\sc owasp}~\cite{OWASPtop10},
{\sc mitre}~\cite{MITREtop25}, and other organizations.
For instance, consider the case of {\sc owasp}'s popular Top Ten
project~\cite{OWASPtop10},
which aims to raise awareness about web application security by
identifying some of the most critical risks organizations may face.
In its three consecutive Top Ten lists (2007, 2010, 2013), different source
code-driven injection attacks dominate the top five positions.

At the same time, attackers find new ways~\cite{HNSHS12,DKH14}
to bypass defense mechanisms using a variety of techniques,
despite the numerous countermeasures that are being introduced.
As an example, already by 2006,
there were more than 20 proposed defenses
against {\sc sql} injection attacks~\cite{HVO06}.
Since then, the number has doubled, while researchers have indicated that
the number of {\sc sql} injection attacks has been steadily
increasing in recent years~\cite{SSL12}.

In this paper, we explore how different attacks associated with 
the exploitation of untrusted input validation errors can be modeled 
under a {\it common perspective}. To that end, we propose an exploitation
model which highlights that most of the steps needed to mount
different types of source code injection attacks are
common.\footnote{Note that we do not consider lower-level attacks based on
the exploitation of memory corruption vulnerabilities and the injection
of binary code---see Section~\ref{sec:attacks}.} This is validated by the fact that
there are protection mechanisms that can defend against more than one of these
types of attacks. Then, we {\it categorize} a selection of representative
protection mechanisms and finally we {\it analyze} each mechanism across the
following dimensions:
\begin{itemize}
\item {\it Accuracy:} protection mechanisms are as good
  as their detection capability; this requires low false positive and
  false negative rates.
\item {\it Availability:} whether the mechanism and its
  testbed are publicly available.
\item {\it Computational performance:} the computational
  overhead of the mechanisms at their points of deployment.
\item {\it Ease of use:} whether the protection
  mechanism is practical in terms of deployment
  and can be easily adopted by security experts.
\item {\it Security:} the robustness of the protection mechanism against
  attackers with knowledge of its internals who attempt to circumvent it.
\item {\it Detection Point:} the location where a mechanism detects an attack
  based on our exploitation model.
\end{itemize}

@@

The rest of the paper is organized as follows.
Section~\ref{sec:attacks} provides some insights on
source code injection attacks and
Section~\ref{sec:model} presents our proposed model.
Section~\ref{sec:dimensions} introduces the dimensions
across which we analyze the various defenses.
Our categorization and analysis is presented in
Section~\ref{sec:defs} and a discussion is provided in
Section~\ref{sec:discussion}. Finally,
Section~\ref{sec:lessons-learned} highlights
some lessons learned from our observations
and Section~\ref{sec:conclusion} concludes the paper.

\section{Source Code Injection on Web Applications}
\label{sec:attacks}

@@

Figure~\ref{fig:taxonomy} presents a taxonomy of source
code injection attacks. Such attacks may involve source code, either of a
{\it Domain Specific Language} ({\sc dsl}) or a {\it Dynamic Language}.
For both categories, we provide examples that will be used
throughout the whole paper.

Application attacks that involve {\sc dsl}s constitute an important
subset of the code injection problem, as {\sc dsl}s like {\sc sql} and
{\sc xml} play an significant role in the development of both web and
mobile applications. For example, many applications have interfaces
where a user enters input to interact with the application's data,
thereby interacting with the underlying database. This input can
become part of an {\sc sql} query and get executed on the target
database. A code injection attack that exploits the vulnerabilities of
these interfaces by taking advantage of input validation issues like
incorrectly passed parameters or incorrect type handling is called an
{\sc sql} injection attack~\cite{CERT02,MS09,HVO06,SW06}.
Consider a trivial example that takes advantage of incorrectly
filtered quotation characters. In an application that
will send the password for a forgetful user via email by executing the
following query:

\bgroup \lstset{language=SQL}
\begin{small}
\begin{lstlisting}
SELECT password from userdata WHERE id = 'Alice'
\end{lstlisting}
\end{small}
\egroup

\noindent
If an attacker, inputs the string {\tt anything' OR 'x'='x}, he or
she could could obtain the administrator's password by email.
Savvy programmers could use a language's libraries, like {\sc php}'s
{\tt mysql\_real\_escape\_string()} to detect malformed input; or they
could use prepared {\sc sql} statements, instead of statement
templates. Unfortunately,
the increasing number of {\sc sql} injection attacks
suggest that programmers are not always that careful.
By using similar techniques malicious users can perform other exploits based on
{\sc dsl}s like {\sc xp}ath~\cite{SW06,CDL07,MKS09},
{\sc xml}~\cite{MSM13} and {\sc json}~\cite{SMS13}. With this kind of
attacks, a malicious user can view sensitive information, destroy or
modify protected data, or even crash the entire application.

{\sc html} is another {\sc dsl} that can be used for malicious purposes
when an application does not properly handle user supplied data. Based
on this vulnerability an attacker can supply valid {\sc html},
typically via a parameter value, and inject their own content into the
page. {\sc html} injection is mainly associated with {\sc xss}
attacks~\cite{BJM08,LZRL09}. However, it can also be used as a
vehicle for an {\sc csrf} attack in the following manner:
Consider a vulnerable bulletin board system where {\tt img} tags are not checked.
A malicious user could embed a {\sc csrf} request within {\tt img}
tags in the following manner:

\bgroup \lstset{language=HTML}
\begin{small}
\begin{lstlisting}
<img src='http://www.vulnerable.com/
  admin.php?edituserwithID=13&addgroup=admin'/>
\end{lstlisting}
\end{small}
\egroup

\noindent
When the page with this injected code is accessed by an administrator,
the attacker (with {\sc id} 13) will gain administrative privileges
while the administrator will have no immediate indication that there
has been an attack.

A recent class of code injection attacks involve dynamic languages
like JavaScript and {\sc php}~\cite{SFVM09,EWKK09,SMS13}.
A simple example of a dynamic language-driven attack is an input string
that is fed into an {\tt eval()} function call, e.g., in {\sc php}:

\bgroup \lstset{language=PHP}
\begin{small}
\begin{lstlisting}
$variable = $_GET['var']; 
$input = $_GET['value'];
eval('$variable = ' . $input . ';')
\end{lstlisting}
\end{small}
\egroup

\noindent
The user may pass into the {\tt value} parameter code that will
execute in the server. Hence, if an attacker provides as
input the following string: {\tt 10 ; system(''touch foo'');}
then a file will be created on the server; it is easy
to imagine more detrimental instances.

JavaScript injection attacks, make up a large subset of dynamic
language-driven attacks. Such attacks are manifested when a web
application accepts and redisplays data of uncertain origin without
appropriate validation and filtering. Based on this flaw, an attacker
can manage to inject a script in the JavaScript engine of a browser
and alter its execution flow~\cite{ELX07}. For instance,
consider a web page (foo.net) that prints the value
of a query parameter ({\tt query}) from the
page's {\sc url} in the web page's content
without escaping the value. A malicious user
can take advantage of this and inject an {\tt iframe} tag
into the page to steal the user's cookie and
send it via an image request to a web site
that he or she controls (malicious.com).
This could be achieved by including the following
link to the malicious web site (or send it via phishing
email) and induce the user to click on it:

\noindent
JavaScript injection is considered a critical issue in web application security
mainly because it is associated with major vulnerabilities such as
{\sc xss} attacks~\cite{SG07} and Cross-Channel Scripting
({\sc xcs}) attacks~\cite{W10,BBB09}. In many cases {\sc xss} attacks
involve the injection of both {\sc html} and JavaScript code.

\section{Exploitation Model}
\label{sec:model}

In this section we provide a step-by-step exploitation model to
understand the process of carrying out application attacks associated
with source code injection. Figure~\ref{fig:attacks}
presents the steps of different application attacks. Links between
steps are labeled with the paths constituting different attacks. The
sequence of steps for each attack is shown by the numbers next to the
attack names in the labels. Most steps are common in all attacks. This
is because they are based on similar attack vectors as we mentioned in
the previous section. Large crosses on transitions in
Figure~\ref{fig:attacks} indicate the points where defenses detect
attacks.

@@

Once the injected code reaches the vulnerable application, it becomes
a part of a value represented by a program variable. The target of the
attack determines the route from now on. In an {\sc sql} injection
attack the injected code becomes part of a query that
finally reaches the database where it is executed.

@@

\begin{figure*}
\begin{center}
\leavevmode
\includegraphics[scale=0.61]{attacks-steps-new.pdf}
\end{center}
\caption{\label{fig:attacks}Attack model demonstrating web
application attacks based on source code injection.
For every arrow there are corresponding transition numbers for
every attack. Every {\it point} indicates a transition where mechanisms
detect attacks. For every point we provide the corresponding
references:
Point {\sc ub} (at the user's browser):~\cite{KJKV09,LV09,TNH07,NSS06,APKLM10,ML10,YCIS07,PSC09,VDDPJ11,OWVS08,DDHPJ10,VFJKKV07,SLMS14,BV08},
Point {\sc s}t{\sc b} (en route from the server to the browser):~\cite{RDWDE07,JKK06a,GC09,JB07,NLC07,WPLKK09,JEP08,PS11},
Point {\sc e}i{\sc q} (embedded in a query en route to the database):~\cite{BWS05,SW06,HCF05,XBS06,PB05,PMP11,MS09,HO05,SMS13},
Point i{\sc db} (in the database):~\cite{BK04,LLW02,VMV05}.}
\end{figure*}

@@

To ensure replicability of research results the source code
implementing a detection mechanism should be available to researchers.
Ideally the code should be available under an open source license, so
that it is easy to modify and improve an approach; even when this is
not possible, for whatever reason, it is important to make sure that
all computer code and test data is available noting any restrictions
on accessibility. Availability of computer code has been recognized as
an important issue outside the computer science field: the journal
\emph{Nature}, in a recent editorial, adopted a publication policy stressing
access to code and data~\cite{nature2014}. At a time and date that the
merits of open access to code are discussed in non computer science
journals~\cite{easterbrook2014}, we should expect computer scientists
to lead the way.

@@

\begin{figure*} [ht]
\begin{center}
\leavevmode
\includegraphics[scale=0.65]{defenses.pdf}
\end{center}
\caption{\label{fig:defenses}The basic categories of countermeasures
against web application attacks based on source code
injection. For each approach we provide the references
that present corresponding mechanisms:
{\it Policy Enforcement}:~\cite{NSS06,JKK06a,KKVJ06,KJKV09,TNH07,RDWDE07,YCIS07,OWVS08,PSC09,ML10,DDHPJ10,PS11,VDDPJ11,BV08},
{\sc isr}:~\cite{BK04,JB07,GC09,APKLM10},
{\it Parse-Tree Validation}:~\cite{BWS05,SW06},
{\it Taint Tracking}:~\cite{HCF05,PB05,XBS06,NLC07,VFJKKV07,PMP11,SLMS14},
{\it Training}:~\cite{LLW02,HO05,HO06,HO05b,VMV05,JEP08,WPLKK09,MS09,MKS09,MKLS11},
{\it Hybrid}:~\cite{BV08,LV09,SMS13}.}
\end{figure*}

We categorize and describe the various mechanisms developed to detect
the attacks described in Section~\ref{sec:attacks} and~\ref{sec:model},
on the dimensions we proposed apart from availability, which we treat
separately in the discussion in Section (\ref{sec:discussion}).
All mechanisms have been proposed in publications cited more than
20 times. We also include some recent mechanisms that have been
presented in top security conferences, even though they have been
cited less than twenty times, so that recent research is not penalized.

Figure~\ref{fig:defenses} presents a taxonomy of the
defenses that protect web applications against
source code injection attacks based on.
We can identify three approaches:
{\it etiological}, {\it symptomatic}, and {\it hybrid}.
The etiological category involves mechanisms designed to
stop attacks based on their causes and origins~\cite{JL75,L81}. 
The symptomatic category incorporates a variety of schemes that
inspect the behavior of applications and detect attacks based on
their undesirable symptoms~\cite{D76,A00}.
Hybrid mechanisms borrow characteristics from both
categories. Table~\ref{tab:comp} groups the various subcategories and
for every mechanism provides the following:

@@

A number of the frameworks define policies based on information and
features provided by the Document Object Model ({\sc dom}) of a web
page. Specifically, developers must place all legitimate scripts
inside {\sc html} elements like {\tt div}. The web browser (Point {\sc
  ub}), parses the {\sc dom} tree and executes scripts only when they
are contained in such elements. All other scripts are treated
according to the policies defined on the server. Frameworks that
support this functionality are {\sc beep}~\cite{TNH07} and {\sc
  dsi}~\cite{NSS06}. The main problem with these mechanisms is that
they do not examine the script's location inside the web document.
Based on this fact, attackers can perform mimicry attacks~\cite{WS02}.
Specifically, they can execute legitimate scripts, but not as intended
by the original design of the developers. This is extensively
described by {\it Athanasopoulos et al.}~\cite{APKLM10}, who also describe
another recent variation of JavaScript injection attacks, known as
return-to-JavaScript attacks that can be used to bypass the above
mechanisms.

@@

An issue regarding the above frameworks
involves features like script inclusion
and {\tt iframe} tags. Even though they allow developers
decide if they will disable them or not,
this is unpractical because such features are quite popular.
If developers choose to use them, these frameworks cannot
define policies that restrict the behavior of third-party
scripts introduced by such features. Thus, they would
be vulnerable in an attack that utilizes {\tt iframes}
in the way we described in Section~\ref{sec:attacks}.
\noindent {\it WebJail}~\cite{VDDPJ11}, and {\sc soma}~\cite{OWVS08}
are two frameworks that can actually detect such an attack (Point {\sc
  ub}). To achieve this, {\sc soma} requires site administrators to
specify legitimate, external domains for sending or receiving
information in order to approve interactions between them and the
protected web site. As a result, {\sc soma} can also detect {\sc csrf}
attacks. WebJail contains the functionality of third-party scripts by
introducing a web component integrator that restricts the access that
these scripts may have to either the data or the functionality of
other components.

@@

Defenses based on training include some mechanisms that can be easily
circumvented. For example, {\sc didafit} and {\it libAnomaly} do not
tag transactions with their corresponding call sites. This can lead
easily to false negatives. For instance, the application
mentioned in Section~\ref{sec:attacks}, which will send the password
for a forgetful user via email by executing the following query:

\bgroup \lstset{language=SQL}
\begin{small}
\begin{lstlisting}
SELECT password from userdata WHERE id = 'Alice'
\end{lstlisting}
\end{small}
\egroup

\noindent
This same application could allow users to lock their terminal,
but allow the unlocking either with the user's password or with
the administrator password (the 4.3 {\sc bsd} {\em lock} command
behaved in this peculiar way).
The corresponding query to verify the password on the locked
workstation would be as follows:

\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
SELECT password from userdata WHERE id = 'Alice' OR
id = 'admin'
\end{lstlisting}
\end{small}
\egroup

\noindent
Even if the application's administrators have chosen to use
either {\sc didafit} or {\it libAnomaly}, an attacker could bypass
them and obtain the administrator's password via email,
by entering on the form the standard string:
{\tt nosuchuser' OR id = 'admin}. The infected query matches the
signature of the second one above and is therefore accepted.
The problem lies with the call location of the query, not the query alone.

@@

\section{Observations}
\label{sec:discussion}

In this section we enumerate key observations
based on the dimensions we set in Section~\ref{sec:dimensions}.

@@

Regarding the subjects of tests, we see that 27 out of 34 mechanisms
were tested on real-world applications known to be vulnerable. The
vulnerabilities associated with these applications are enlisted in the
following providers: the Common Vulnerabilities and Exposures ({\sc
  cve}) database~\cite{cve}, the {\it Bugtraq} security
mailing list~\cite{bugtraq}, the {\sc xss}ed.com
security bulletin provider~\cite{xssed}, Microsoft's security
bulletin~\cite{microsoftBulletin} and the
ha.ckers.org security bulletin provider~\cite{hackers}.
Figure~\ref{fig:defect_sources} presents how many, in percentage,
and which, in the figure's caption, authors referred
to which source. In numerous occassions authors performed tests based
on the same applications. For example, 12 mechanisms were tested on a
vulnerable version of {\sc phpbb}.\footnote{https://www.phpbb.com/}

@@

\section{Recommendations and Lessons Learned}
\label{sec:lessons-learned}

In this section we provide some lessons learned from
our observations and provide some recommendations
that developers of new mechanisms may find helpful.

@@

\subsection{Improving Accuracy Tests}

@@

When we introduced specificity and sensitivity in
Section~\ref{ssec:diagnostic-performance} we deferred discussion of
\textsc{ppv} and \textsc{npv} to this point. These two relate to the
performance of a detection mechanism in an actual production setting,
instead of a testbed. If an attack is detected in a production
environment, how much should we be worried? The answer is provided by
\textsc{ppv}. If no attack is detected in a production environment,
how relaxed should we be that no attack has indeed taken place? The
answer is provided by \textsc{npv}. We can calculate \textsc{ppv} and
\textsc{npv} with the following equations:

@@

\subsection{On Code Availability}

Apart from testing practices, another area that merits improvement is
the availability of source code and testbeds. We are not aware of
specific reasons why authors of attack detection mechanisms seem to be
wary of releasing on-line their code and tests. Our finding may reflect
the status in the current point in time, when authors are urged to
publish their code and tests, and may start, or may have already
started doing so, but this does not show yet in the research we
examined, which goes several years back. We also saw instances where
material was published, but does not seem to be available any more.
That points to the importance of curation of research materials. That
is a problem arising in all scientific fields. It is not enough to
publish the underlying code and data, but to make sure that it remains
accessible and to provide the means to test it even as technology
progresses and operating systems, file formats, and software libraries
change. That may be too much to ask right now; what seems reasonable,
though, is to ask of researchers in attack detection mechanisms not to
buck the trend and to take steps to increase the availability of their
research.

\subsection{Computational Overhead and Deployment Remarks}

@@

@@

Despite many approaches that have been developed,
application attacks based on source code injection
have been consistently present for the last 15 years,
and it appears that they will continue to be.
Attackers seem to find new ways to introduce
malicious source code to applications by using a
variety of languages and techniques~\cite{HNSHS12,DKH14}.
Meanwhile, during the last decade, there are
numerous mechanisms that have been developed to
detect one or more of such attacks. However,
even if some of them share characteristics
(for instance, {\sc html} sanitization and {\tt eval} handling)
with supported frameworks like Mozilla's Content Security
Policy~\cite{cspMoz} ({\sc csp}) and Google Caja~\cite{M06,caja}
the majority of them are not used in practice.
